---
title: "Assessment1"
author: "Laurig Gomez - s4635952"
date: "19/12/2021"
output: html_document
---

# **R Markdown editing**

*R Markdown is a file format for making dynamic documents with R. An R markdown document is written in markdown ( an easy-to-write plain text format) and contains chunks of embedded R code.*
*For more details of R Markdown see* *[Here](https://rmarkdown.rstudio.com/lesson-8.htm)*.

6. Insert an R chunk and create a variable with this dataset
https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv
The name of the variable should be `fastfood`

```{r}
library(tidyverse)
library(knitr)
fastfood<-
  read.csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')
```


```{r}
head(fastfood,10)%>%
  kable()
```

Display the first 10 rows of the dataset and the first 5 variables
```{r}
kable(fastfood[1:10.1:5])
```

8.Display the observations that has more than 1000 calories
```{r}
filter(fastfood,calories>1000)%>%
  kable()
```

9. Arrange observations with more than 40 in total_fat and more than 80 in total_carb in the descending order and save them to a new variable (dataset) called `dont_eat_this`
```{r}
#Arrange
dont_eat_this<-fastfood %>% filter(total_fat>40 & total_carb>80, decreacing=TRUE)
#display
fastfood%>%filter(total_carb>80)%>% kable()
```


10. Using the initial dataset variable, use `case_when()` to create a new variable `heavy_food` which is equal to "heavy" when total calories are greater than 500, "low" when total calories are less than 250 and "average" for all other cases. Count the number of observations in each created category.
```{r}
```


11. Display the types of variables in the dataset using `skimr` package

```{r}
library(skimr)
skim(fastfood)
```

12. Present the count observations from each restaurant in a descending order
```{r}
fastfood%>%
  count(restaurant,wt=NULL,sort=TRUE,name = "res_num",decreasing=TRUE)%>%
  ungroup()%>%
  kable()
```


13. Using groupings (group_by()), summarise and display the average number of calories for each restaurant.
```{r}
fastfood%>%
  group_by(restaurant)%>%
  summarise(average_cal = mean(calories, trim = 0, na.rm = FALSE)) %>%
  ungroup() %>%
  kable()
```

14. Add variables to the dataset, which:

-  calculates the average calories per type of restaurant and call it `average_calories`
```{r}
average_calories <-
  fastfood %>%
  group_by(restaurant)%>%
  mutate(average_calories = mean(calories)) %>%
  ungroup()
head(average_calories %>% select(restaurant,calories,average_calories))
```

-  calculates the maximum total_fat per type of restaurant and call it `max_fat`
```{r}
max_fat<-
  fastfood %>%
  group_by(restaurant)%>%
  mutate(max_fat = sum(cal_fat,sat_fat, trans_fat)) %>%
  ungroup()
head(max_fat %>% select(restaurant,calories,total_fat))
```

-  calculates the minimum cholesterol per type of restaurant and call it `min_cholesterol`
```{r}
min_cholesterol <-
  fastfood %>%
  group_by(restaurant)%>%
  mutate(min_cholesterol = min(cholesterol)) %>%
  ungroup()
head(min_cholesterol%>% select(restaurant,cholesterol,min_cholesterol))
```


15. Display the data vis of total fat per each type of restaurant. Write a narration (2-3 sentences) why you believe this type of data viz presents such information best.
```{r}
### This question would be found two different kind of variables: one would be displayed as Restaurant and the other one would be clasify as total_fat). Normally to  make a ggplot2 or chart it would be necessary to identify variable like if it is discrete or continuous, both continuos or both discrete;so, for restaurant would be identify as discrete and about total fat  would be a continuous. Bar chart would be the easiest one to understand when the difference between each restaurant.


```{r}
# variable_x = restaurant (this is discrete) and variable_y = total_fat (this is continuous)
fastfood %>%
  group_by(restaurant) %>%
  ungroup() %>%
  ggplot(aes(x = restaurant, y = total_fat)) +
  geom_col() +
  labs(x="Names of Restaurant", y="Total Fat", title="Data Visualisation 1", caption = "Image 1:Bar Graph")
```

16. Add a variable to the dataset, which calculates the sum of cholesterol and sodium and call it `cholesterol_sodium`.

Remove the variable `salad`

17. Use observations for McDonald to plot sugar variable against protein with `geom_point()`
```{r}
fastfood %>%
  select(restaurant, sugar, protein) %>%
  filter(restaurant=="Mcdonalds") %>%
  ggplot(aes(sugar,protein)) +
  geom_point(colour = "blue") +
  labs(x="Sugar", y="Protein", title = "Data Visualisation - Mcdonalds", caption = "Image 2:Point Graph")
```


**Focus**: ggplot2, factors, strings, dates  

18. Identify variable(s) which should be factors and transform their type into a factor variable.

```{r}
#install.packages("forcats")
library(forcats)
#is.factor(fastfood$restaurant)
#as.factor(fastfood$restaurant)
```

19. Create two new variables:

Read about `cut_number()` function using Help and add a new variable to the dataset `calories_type`. Use `calories` variable for `cut_number()` function to split it into 3 categories `n=3`, add labels `labels=c("low", "med", "high")` and make the dataset ordered by arranging it according to calories.


Do not forget to save the updated dataset.

```{r}
calories_type <- fastfood %>%
  group_by(restaurant) %>%
  mutate(calories_type = cut_number(calories, n = 3, labels = c("low","med","high"))) %>%
  ungroup() %>%
  arrange(calories_type)
head(calories_type %>% select(restaurant,item,calories,calories_type))
```

20. Create a dataviz that shows the distribution of `calories_type` in food items for each type of restaurant. Think carefully about the choice of data viz. Use facets, coordinates and theme layers to make your data viz visually appealing and meaningful. Use factors related data viz functions.

```{r}
calories_type %>%
  group_by(restaurant) %>%
  ungroup() %>%
  ggplot(aes(restaurant, calories_type)) +
  geom_boxplot(colour="red") +
  facet_wrap(vars(restaurant)) +
  coord_flip() +
  theme_bw() +
  labs(x="Restaurant", y="Calories Type", title ="Data Visualisation 3", caption = "Image 3:Plot Graph")
 
```

21. The second variable should show the percentage of `trans_fat` in `total_fat`. Add the variable to the dataset and call it `trans_fat_percent`. Do not forget to save the updated dataset.
```{r}
trans_fat_percent <- fastfood %>%
  group_by(restaurant) %>%
  mutate(trans_fat_percent = trans_fat / total_fat * 100) %>%
  ungroup()
head(trans_fat_percent %>% select(restaurant,item,trans_fat,total_fat,trans_fat_percent))
#round(trans_fat_percent$trans_fat_percent, 2)
```

22. Create a dataviz that shows the distribution of `trans_fat` in food items for each type of restaurant. Think carefully about the choice of data viz. Use facets, coordinates and theme layers to make your data viz visually appealing and meaningful.
```{r}
ggplot(fastfood, aes(restaurant,trans_fat)) +
  geom_point(colour = "yellow") +
  facet_wrap(vars(restaurant)) +
  coord_flip() +
  labs(x="Restaurant", y="Trans Fat", title = "Data Visualisation 4", caption = "Image 4:Point Graph") +
  theme_dark()
```

23. Calculate and show the average (mean)  `total_fat` for each type of restaurant. No need to save it as a variable.
```{r}
fastfood %>%
  group_by(restaurant) %>%
  summarise(average_fat = mean(total_fat)) %>%
  ungroup() %>%
  kable()
```


24. And create a dataviz that allow to compare different restaurants on this variable (`total_fat`). You can present it on one dataviz (= no facets).

```{r}
#ggplot(fastfood, aes(restaurant,total_fat)) + geom_col()
fastfood %>%
  group_by(restaurant) %>%
  summarise(average_fat = mean(total_fat)) %>%
  ungroup() %>%
  mutate(average_fat = fct_reorder(restaurant,average_fat)) %>%
  ggplot(aes(restaurant,average_fat)) +
  geom_col()
```